---
title: "p8105_hw3_yc4384"
author: "Yangyang Chen"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem_1

**1. Importing Data**
```{r, echo=FALSE, message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
data("instacart")
colnames(instacart)
```

**2. Data Description**

* The dataset **instacart** is a `r nrow(instacart)` * `r ncol(instacart)` dataframe. It contains `r nrow(instacart)` observations and key variables are: `r colnames(instacart)[7:15]`. Dataset has the order records of each costumers, including _product name, order number, user id_, etc.


* There are **134** aisles, and  **fresh vegetables** aisles are the most items ordered from:
```{r, message=FALSE}
instacart |> 
  group_by(aisle) |> 
  summarize(n = n()) |> 
  count()
```

```{r, message=FALSE}
instacart |> 
  group_by(aisle) |>
  summarize(n = n()) |> 
  mutate(n_ranking = min_rank(n)) |> 
  filter(n_ranking>133) 
```

* Following plot shows **the number of items ordered in each aisle**, with limitation that to aisles with more than 10000 items ordered:
```{r, message=FALSE}
instacart |> 
  group_by(aisle) |>
  summarize(n = n()) |> 
  mutate(n_ranking = min_rank(n)) |> 
  filter(n>10000) |> 
  ggplot(aes(x = n, fill = aisle)) +
  geom_histogram()
```

* Following table shows **the three most popular items in each of the aisles** “baking ingredients”, “dog food care”, and “packaged vegetables fruits”:

```{r, message=FALSE}
instacart |> 
  group_by(aisle, product_name) |> 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") |> 
  summarize(n = n()) |> 
  filter(min_rank(desc(n)) < 4) |> 
  knitr::kable()
```

* Following table shows the mean hour of the day** at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r, message=FALSE}
instacart |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream" ) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable()
```

## Problem 2
**1. Importing Data**
```{r, message=FALSE}
data("brfss_smart2010")
```

**2. Data cleaning:**

* Format the data to use appropriate variable names;

* Focus on the “Overall Health” topic;

* Include only responses from “Excellent” to “Poor”;

* Organize responses as a factor taking levels ordered from “Poor” to “Excellent”.
```{r, results='hide', message=FALSE}
brfss_clean_df = 
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |> 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor")))

```

* In **2002**, states **`MA, NJ and PA`** were observed at 7 or more locations:
```{r, message=FALSE}
brfss_clean_df |> 
  filter(year==2002) |> 
  group_by(locationabbr) |> 
  distinct(locationdesc) |> 
  summarize(n = n())|> 
  filter(n>7) 
```
* In **2010**, states **`CA, FL, MA, MD, NC, NE, NJ, NY, OH, TX and WA`** were observed at 7 or more locations:
```{r, message=FALSE}
brfss_clean_df |> 
  filter(year==2010) |> 
  group_by(locationabbr) |> 
  distinct(locationdesc) |> 
  summarize(n = n())|> 
  filter(n>7) 
```


**brfss_exce_df:** dataset only includes Excellent responses, and contains year, state, and a variable that averages the data_value across locations within a state:

```{r, message=FALSE}
brfss_exce_df =
  brfss_clean_df |> 
  filter(response == "Excellent") |> 
  select(c(year, locationabbr, locationdesc, data_value)) |> 
  group_by(locationabbr, year) |> 
  mutate(data_value_mean = mean(data_value, na.rm = TRUE))
```

**Spaghetti plot:** average value over time within a state:
```{r, message=FALSE}
brfss_exce_df |> 
  ggplot(aes(x = year, y = data_value_mean, group = locationabbr, color = locationabbr)) +
  geom_line() 
  
```

**Two-panel plot** shows for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State:
```{r, message=FALSE}

brfss_p_exce_df = 
  brfss_clean_df |> 
    filter(response %in% c("Excellent", "Poor")) |> 
    filter(year %in% c(2006, 2010)) |> 
    filter(locationabbr == "NY") |> 
    select(c(year, locationabbr, locationdesc, data_value, response)) |> 
    group_by(locationabbr, year) 

brfss_p_exce_df |>   
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point() +
  facet_grid(. ~ year)
```

## Problem 3

**1. Data cleaning and merging:**

* Dataset includes all originally observed variables; 
* Exclude participants less than 21 years of age, and those with missing demographic data; 
* Encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).
```{r, message=FALSE}
acc_df = 
  read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names()
cov_df = 
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names()
cov_clean_df =
  cov_df |>
  filter(age >= 21) |> 
  drop_na() |> 
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"),
    sex = as.factor(sex)) |> 
  mutate(
    education = 
      case_match(
        education, 
        1 ~ "Low", 
        2 ~ "Median",
        3 ~ "High"),
    education = as.factor(education)) 

mims_data = 
  left_join(cov_clean_df, acc_df, by = "seqn") 
```
**2. Data Visualization**

**Table**: the number of men and women in each education category.
```{r, message=FALSE}
mims_data |> 
  group_by(sex, education) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = sex,
    values_from = n
  ) |> 
  knitr::kable()
```
**Comments:**

* More females have received high and low level education than male.

* More male have received median level education than female.

**Plot**: Create a visualization of the age distributions for men and women in each education category.
```{r, message=FALSE}
mims_data |> 
  group_by(sex, education, age) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = age, fill = education)) +
  geom_density(alpha = .4, adjust = .5, color = "blue") +
  facet_wrap(~ sex, nrow = 1)
```
**Comments:** 

* We can observe that female group received high education in age 60 - 80; While male received more low education than female. 
* Besides, female in each age interval received balanced education levels, and reaches education peak in high education; However, male reaches education peak in low level.


**Three Panel Plot:**

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r, message=FALSE}
mims_data |> 
  select(seqn, age, education, sex, min1:min1440) |> 
    pivot_longer(
      seqn,
      names_to = "total_activities", 
      values_to = min1:min1440
    ) |> 
  ggplot(aes(x = age, y = total_activities, color = sex)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom") +
  facet_wrap(~ education, nrow = 1)
```

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

