---
title: "p8105_hw3_yc4384"
author: "Yangyang Chen"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem_1

**1. Importing Data**
```{r, echo=FALSE, message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
data("instacart")
colnames(instacart)
```

**2. Data Description**

* The dataset **instacart** is a `r nrow(instacart)` * `r ncol(instacart)` dataframe. It contains `r nrow(instacart)` observations and key variables are: `r colnames(instacart)[7:15]`. Dataset has the order records of each costumers, including _product name, order number, user id_, etc.


* There are **134** aisles, and  **fresh vegetables** aisles are the most items ordered from.
```{r}
instacart |> 
  group_by(aisle) |> 
  summarize(n = n()) |> 
  count()
```

```{r}
instacart |> 
  group_by(aisle) |>
  summarize(n = n()) |> 
  mutate(n_ranking = min_rank(n)) |> 
  filter(n_ranking>133) 
```

* This plot shows **the number of items ordered in each aisle**, with limitation that to aisles with more than 10000 items ordered. 
```{r}
instacart |> 
  group_by(aisle) |>
  summarize(n = n()) |> 
  mutate(n_ranking = min_rank(n)) |> 
  filter(n>10000) |> 
  ggplot(aes(x = n, fill = aisle)) +
  geom_histogram()
```

* This table shows **the three most popular items in each of the aisles** “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart |> 
  group_by(aisle, product_name) |> 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") |> 
  summarize(n = n()) |> 
  filter(min_rank(desc(n)) < 4) |> 
  knitr::kable()
```

* This table shows **the mean hour of the day** at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream" ) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable()
```

## Problem 2
**1. Importing Data**
```{r}
data("brfss_smart2010")
```

**2. Data cleaning:**

* Format the data to use appropriate variable names;

* Focus on the “Overall Health” topic;

* Include only responses from “Excellent” to “Poor”;

* Organize responses as a factor taking levels ordered from “Poor” to “Excellent”.
```{r, results='hide'}
brfss_clean_df = 
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |> 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor")))

```

* In **2002**, states **`MA, NJ and PA`** were observed at 7 or more locations:
```{r}
brfss_clean_df |> 
  filter(year==2002) |> 
  group_by(locationabbr) |> 
  distinct(locationdesc) |> 
  summarize(n = n())|> 
  filter(n>7) 
```
* In **2010**, states **`CA, FL, MA, MD, NC, NE, NJ, NY, OH, TX and WA`** were observed at 7 or more locations:
```{r}
brfss_clean_df |> 
  filter(year==2010) |> 
  group_by(locationabbr) |> 
  distinct(locationdesc) |> 
  summarize(n = n())|> 
  filter(n>7) 
```


* **brfss_exce_df** dataset only includes Excellent responses, and contains year, state, and a variable that averages the data_value across locations within a state. 

```{r}
brfss_exce_df =
  brfss_clean_df |> 
  filter(response == "Excellent") |> 
  select(c(year, locationabbr, locationdesc, data_value)) |> 
  group_by(locationabbr, year) |> 
  mutate(data_value_mean = mean(data_value, na.rm = TRUE))
```

* Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss_exce_df |> 
  ggplot(aes(x = year, y = data_value_mean, group = locationabbr, color = locationabbr)) +
  geom_line() 
  
```

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}

brfss_p_exce_df = 
  brfss_clean_df |> 
    filter(response %in% c("Excellent", "Poor")) |> 
    filter(year %in% c(2006, 2010)) |> 
    filter(locationabbr == "NY") |> 
    select(c(year, locationabbr, locationdesc, data_value, response)) |> 
    group_by(locationabbr, year) 

brfss_p_exce_df |>   
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point() +
  facet_grid(. ~ year)
```

## Problem 3
Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices can measure MIMS in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses accelerometer data collected on 250 participants in the NHANES study. The participants’ demographic data can be downloaded here, and their accelerometer data can be downloaded here. Variables *MIMS are the MIMS values for each minute of a 24-hour day starting at midnight.

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

